On the Gravitational stability of gravito-turbulent accretion disks

### 2016-01-31 paper/summary.tex
minor edit
minor summary edits, new alpha(gamma) relation based on gammie's alpha and our betastar

\subsection{Relation to numerical simulations}\label{prev_works}
%frag in sims because of : cooling, viscosity, or high den
%flucuations? don't know 
Our results may help understand some numerical simulations 
concerning disk fragmentation. Table 
\ref{bstar_compare} shows a close match between the characteristic
cooling time for cooling-driven GI (Eq. \ref{betastar}) 
and that for disk fragmentation observed in simulations 
\citep{gammie01,rice05,rice11}. This suggests that, at least for those
simulations, fragmentation is physically due to the removal of
thermal stabilization by cooling on radial lengthscales of the disk 
thickness.   
%i.e. there is no effective heating mechanism on the scale of
        %H 
%{\bf This would imply over-densities on this scale 
%do not experience heating during its growth. This is not unreasonable
%because in reality, gravito-turbulent heating only occurs when a 
%spiral shock passes through said over-density \citep{young16}. Thus
%there there is a time interval during which only cooling acts, whence
%our inviscid theory can be applied. --- too much?}

@@

We suggest having a floor temperature in the standard beta
cooling prescription is a necessary for  
numerical convergence. This limits the
relevant scales to a finite range.  %for cool gi 
%physically irradiation 
Furthermore, without a floor temperature, standard beta cooling is
a function of the pressure/energy density field only. It is not clear
how relevant are results obtained from this is to actual PPDs,
where cooling depends on two thermodynamic states (e.g. pressure and
temperature). A floor temperature permits a mapping between standard beta
cooling and PPD cooling (\S\ref{ppd_cooling}). 
%KMK
Moreover, the standard beta cooling does 
not account for optical depth effects, which force the midplane 
and high density perturbations to cool more slowly.

>>>>

\subsection{Relation to numerical simulations}\label{prev_works}
%frag in sims because of : cooling, viscosity, or high den
%flucuations? don't know 
Our results may help understand some numerical simulations 
concerning disk fragmentation. Table 
\ref{bstar_compare} shows a close match between the characteristic
cooling time for cooling-driven GI (Eq. \ref{betastar}) 
and that for disk fragmentation observed in simulations 
\citep{gammie01,rice05,rice11}. This suggests that, at least for those
simulations, fragmentation is physically due to the removal of
thermal stabilization by cooling on radial lengthscales of the disk 
thickness. In this interpretation, gravito-turbulence only
provides a background heating. %not active in frag 
Our characteristic cooling time, given by 
Eq. \ref{betastar}, corresponds to a dimensionless viscosity 
as defined in the above studies\footnote{This differs from our
  definition of $\alpha$ by a factor of $\gamma^{-1}$.}  
\begin{align}
  \alpha =
  \frac{4}{9}\frac{1}{\gamma}\frac{\left(\sqrt{\gamma}-1\right)^{1/2}}{\left(\sqrt{\gamma}+1\right)}
  \simeq \begin{cases}
    0.062 & \gamma = 7/5, \\
    0.063 & \gamma = 5/3,\\
    0.059 & \gamma = 2.
  \end{cases}
\end{align}
%where we have inserted a factor of $\gamma^{-1}$ to fit the
%viscosity definition used by \cite{gammie01}.

%i.e. there is no effective heating mechanism on the scale of
        %H 
%{\bf This would imply over-densities on this scale 
%do not experience heating during its growth. This is not unreasonable
%because in reality, gravito-turbulent heating only occurs when a 
%spiral shock passes through said over-density \citep{young16}. Thus
%there there is a time interval during which only cooling acts, whence
%our inviscid theory can be applied. --- too much?}

@@

We suggest having a floor temperature in the standard beta
cooling prescription is  necessary for  
numerical convergence. This limits the
relevant scales to a finite range.  %for cool gi 
%physically irradiation 
Furthermore, without a floor temperature, standard beta cooling is
a function of the pressure/energy density field only. It is not clear
how relevant are results obtained from this is to actual PPDs,
where cooling depends on two thermodynamic states (e.g. pressure and
temperature). A floor temperature permits a mapping between standard beta
cooling and PPD cooling (\S\ref{ppd_cooling}). 
%KMK
Moreover, the standard beta cooling does 
not account for optical depth effects, which force the midplane 
and high density perturbations to cool more slowly.
